### 2015-07-18 introduction.tex
Change name of tex file to 'spack-paper-sc15'
combinatoric -> combinatorial

%!TEX root = paper.tex

@@

To support the diverse needs of applications, system administrators 
and application developers 
frequently build, install, and support many different configurations
of math and physics libraries, as well as other software.
Frequently, applications must be rebuilt to fix bugs and to support
new versions of the operating system (OS), Message Passing Interface (MPI)
implementation, compilers, and other dependencies. 
Unfortunately, building scientific software is notoriously complex, with
immature build systems that are difficult to adapt to new
machines~\cite{dubois+:comp-sci-eng,hoste+:pyhpc12,wilson+:corr}.
Worse, the space of required builds grows combinatorically
with each new configuration parameter. As a result, LLNL staff
spend countless hours dealing with build and deployment issues.

Existing package management tools automate parts of the build 
process~\cite{bsdports,digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,homebrew,hoste+:pyhpc12,macports,thiruvathukal:gentoo04}.
For the most part, these systems focus on keeping a single, stable set of 
packages up to date, and they do not handle installation of multiple
versions or configurations.  Those that {\it do} handle multiple configurations
typically require that package files be created for each combination of 
options~ \cite{digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,hoste+:pyhpc12}, 
leading to a profusion of files and maintenance issues.
To our knowledge, no existing tool allows new configurations to be composed
and assembled on demand.  Some allow limited forms of 
composition~\cite{hoste+:pyhpc12,dolstra+:icfp08,dolstra+:lisa04}, but their
dependency management is overly rigid, and they burden users with
the combinatoric problems of naming and versioning.

>>>>

%!TEX root = spack-paper-sc15.tex

@@

To support the diverse needs of applications, system administrators
and application developers frequently build, install, and support many
different configurations of math and physics libraries, as well as
other software.  Frequently, applications must be rebuilt to fix bugs
and to support new versions of the operating system (OS), Message
Passing Interface (MPI) implementation, compilers, and other
dependencies.  Unfortunately, building scientific software is
notoriously complex, with immature build systems that are difficult to
adapt to new
machines~\cite{dubois+:comp-sci-eng,hoste+:pyhpc12,wilson+:corr}.
Worse, the space of required builds grows combinatorially with each
new configuration parameter. As a result, LLNL staff spend countless
hours dealing with build and deployment issues.

Existing package management tools automate parts of the build
process~\cite{bsdports,digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,homebrew,hoste+:pyhpc12,macports,thiruvathukal:gentoo04}.
For the most part, these systems focus on keeping a single, stable set of
packages up to date, and they do not handle installation of multiple
versions or configurations.  Those that {\it do} handle multiple configurations
typically require that package files be created for each combination of
options~ \cite{digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,hoste+:pyhpc12},
leading to a profusion of files and maintenance issues.
To our knowledge, no existing tool allows new configurations to be composed
and assembled on demand.  Some allow limited forms of
composition~\cite{hoste+:pyhpc12,dolstra+:icfp08,dolstra+:lisa04}, but their
dependency management is overly rigid, and they burden users with
the combinatorial naming and versioning problems.

### 2015-07-18 motivation.tex
Change name of tex file to 'spack-paper-sc15'
combinatoric -> combinatorial

\paragraph{Meta-build Systems}
{\it Meta-build systems} such as Contractor, WAF, and
MixDown~\cite{amundson:contractor,epperly+:mixdown,epperly+:mixdown-report,nagy:waf} are
related to package managers, but they focus on ensuring that a single
package builds correctly with its dependencies.  MixDown notably provides excellent
features for ensuring consistent compiler flags in a build.
However, these systems do not provide facilities to manage
large package repositories or combinatoric versioning.

\paragraph{Traditional package managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and
APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package only has a single version
and most of the tools install packages in a single, inflexible location.
To install multiple configurations, the user must create a custom, combinatoric
naming scheme to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

@@

\paragraph{Gaps in current common practice}
The cryptographic versioning of Nix and HashDist is very flexible. It versions
the package {\it and} its dependency configuration, and can represent any
configuration. However, users cannot navigate or query the installed software.
The systems are in a sense ``write-only'':
Nix does not even offer a way to ask for a package's dependencies.
%
EasyBuild and Smithy generate environment modules, which allow querying
through their own interface. Naming schemes used in existing module systems, however, cannot
handle combinatoric versions. The Lmod authors call this the
``matrix problem``~\cite{mclay:lmod-tutorial}.

@@

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatoric, and realizing an arbitrary
combination of compiler, MPI version, build options, and dependency versions
requires tediously modifying many package files.
%
Indeed, the number of package files required for existing systems scales with
the number of version {\it combinations}, not the number of packages, which
quickly becomes unmanageable.  As an example, the very flexible EasyBuild
system has over 3,300 files for several permutations of only 600 packages.
Other systems have similar issues. HPC sites need a mechanism to
{\it parameterize} packages so that new builds can be {\it composed} on the fly.

>>>>

\paragraph{Meta-build Systems}
{\it Meta-build systems} such as Contractor, WAF, and
MixDown~\cite{amundson:contractor,epperly+:mixdown,epperly+:mixdown-report,nagy:waf} are
related to package managers, but they focus on ensuring that a single
package builds correctly with its dependencies.  MixDown notably provides excellent
features for ensuring consistent compiler flags in a build.
However, these systems do not provide facilities to manage
large package repositories or combinatorial versioning.

\paragraph{Traditional package managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and
APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package only has a single version
and most of the tools install packages in a single, inflexible location.
To install multiple configurations, the user must create a custom, combinatorial
naming scheme to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

@@

\paragraph{Gaps in current common practice}
The cryptographic versioning of Nix and HashDist is very flexible. It versions
the package {\it and} its dependency configuration, and can represent any
configuration. However, users cannot navigate or query the installed software.
The systems are in a sense ``write-only'':
Nix does not even offer a way to ask for a package's dependencies.
%
EasyBuild and Smithy generate environment modules, which allow querying
through their own interface. Naming schemes used in existing module systems, however, cannot
handle combinatorial versions. The Lmod authors call this the
``matrix problem``~\cite{mclay:lmod-tutorial}.

@@

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatorial, and realizing an arbitrary
combination of compiler, MPI version, build options, and dependency versions
requires tediously modifying many package files.
%
Indeed, the number of package files required for existing systems scales with
the number of version {\it combinations}, not the number of packages, which
quickly becomes unmanageable.  As an example, the very flexible EasyBuild
system has over 3,300 files for several permutations of only 600 packages.
Other systems have similar issues. HPC sites need a mechanism to
{\it parameterize} packages so that new builds can be {\it composed} on the fly.
